Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance

### 2017-07-21 ms_pre_post.Rmd
Remove the bold for changed text

From the perspective of a breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred (i.e., post-copulatory inbreeding avoidance). Evolution of such post-copulatory inbreeding avoidance is likely to be facilitated when females express some degree of polyandry, defined as mating with multiple males during a single reproductive bout [but see @Dougherty2016]. Such polyandry might itself be under **indirect** selection and evolve if it allows females to mate with additional unrelated males following an initial mating with a relative**, or if it allows females to mate with additional unrelated males that would not be available during initial mate choice due to changes in mate availability or relatedness distribution in a population [e.g., @Reid2015; @Duthie]**. Polyandry can thereby facilitate pre-copulatory inbreeding avoidance even without any post-copulatory female choice or otherwise biased fertilisation among sperm provided by females' mates (i.e., under conditions of a 'fair raffle'). <!--- FIXIT: Reid et al. 2015 comes up as 'b', should be 'a' --->

Overall, polyandry can therefore simultaneously allow females to mate with less closely related males and create the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to allow females to avoid inbreeding and thereby increase offspring fitness [@Zeh1997; @Jennions2000; @Tregenza2002]. Resulting indirect selection on polyandry that stems from the increase in offspring fitness could explain evolution of polyandry in cases where multiple mating does not increase female reproductive success, or even reduces reproductive success and therefore imposes direct costs on polyandrous females. However, despite such widely-invoked hypotheses regarding the role of inbreeding depression in driving pre-copulatory and post-copulatory strategies and polyandry, there is surprisingly little theory or modelling that critically examines the conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are predicted to evolve, or that examines how the evolution of such strategies of inbreeding avoidance might feed back to affect underlying evolution of polyandry. **To comprehensively understand evolution of mating strategies given inbreeding depression, it is necessary to consider the fundamental effects of selection on pre-copulatory and post-copulatory inbreeding avoidance and polyandry.**

Despite the paucity of theory, numerous empirical studies on diverse species have tested for, and in some cases found evidence of, female inbreeding avoidance in systems where polyandry also occurs [@Tregenza2002; @Varian-Ramos2012; @Kingma2013; @Arct2015; but see @Reid2015a]. However, few studies have determined whether inbreeding avoidance is enacted through pre-copulatory or post-copulatory mechanisms. Pre-copulatory inbreeding avoidance has been reported in sweet potato weevils [*Cylas formicarius*, @Kuriwada2011], purple-crowned fairy-wrens [*Malurus coronatus*, @Kuriwada2011], and squinting bush brown butterflies [*Bicyclus anynana*, @Fischer2015], while post-copulatory inbreeding avoidance has been reported in, **e.g.,** red junglefowl [*Gallus gallus*, @Pizzari2004] and crickets [*Teleogryllus oceanicus*, @Simmons2006; *Gryllus bimaculatus*, @Bretman2009]. Evidence for both pre-copulatory and post-copulatory inbreeding avoidance is available across different studies of Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. Meanwhile, @Liu2014 found evidence of pre-copulatory but not post-copulatory inbreeding avoidance within a single study of cabbage beetles (*Colaphellus bowringi*). However, @Ala-Honkola2011 and @Tan2012 found no evidence for pre-copulatory or post-copulatory inbreeding avoidance in fruit flies (*Drosophila melanogaster*), respectively, and Reid et al. [-@Reid2014; -@Reid2015] found no net inbreeding avoidance in song sparrows (*Melospiza melodia*) despite strong inbreeding depression and opportunities for both pre-copulatory and post-copulatory inbreeding avoidance. Taken together, these studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance, or lack thereof, occur in nature. However, there is as yet no theory that predicts what combinations of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry should be favoured by selection when all can evolve. Consequently, there is no theory that allows the diversity of observed pre-copulatory and post-copulatory inbreeding avoidance strategies to be interpreted, and no clear hypotheses that could be tested through future empirical studies of individual systems or subsequent comparative analyses.

@Duthie used a genetically-explicit individual-based model to examine the conditions under which polyandry is predicted to evolve due to selection stemming from pre-copulatory inbreeding avoidance in the absence of any post-copulatory inbreeding avoidance. **Even when selection for pre-copulatory inbreeding avoidance was strong and females consequently preferred unrelated mates, selection for polyandry specifically as a mechanism to facilitate this inbreeding avoidance** occurred only under highly restricted conditions, requiring that direct negative selection (i.e., 'costs') on polyandry was weak, that very few males were available for a female's initial mate choice but many were available for additional mate choice(s), or that polyandry was conditionally expressed when a focal female was related to her initial mate [@Duthie]. Without these conditions, **females that engaged in** polyandry tended **to be constrained** to increase rather than decrease **their** overall degree of inbreeding, **ultimately** reduc**ing** offspring fitness. This **constraint** occurred because, once pre-copulatory inbreeding avoidance evolved, polyandrous females had already chosen available unrelated males as their initial mates. Their additional mates, chosen from the remaining male population, were therefore increasingly likely to include relatives. Evolution of polyandry purely to facilitate pre-copulatory inbreeding avoidance, and hence the resulting degree of inbreeding avoidance, was consequently restricted [@Duthie].

@@

Ultimately, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry will **be affected not only by** the magnitudes of positive indirect selection stemming from reduced inbreeding depression in females' offspring, but also **by** the magnitudes of direct negative selection on resulting phenotypes (i.e., the direct fitness costs of expressing each reproductive strategy). **There are multiple mechanisms by which direct costs might by incurred for polyandry, pre-copulatory choice, and post-copulatory choice; here we focus on three examples observed in empirical populations.** Polyandrous females **might** pay costs of mate searching or mating because these activities increase predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice and consequent reluctance to mate with relatives might pay costs of increased risk of complete mating failure or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. **Finally,** females that express post-copulatory choice might pay initial energetic costs associated with developing the physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. If the relative costs of pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly strategy might become fixed over generations, while alleles underlying the more costly strategy might become extinct, especially if their effects are made redundant by evolution of the less costly strategy. Evolutionary dynamics will therefore depend on the benefits of each strategy, which might also differ and vary in relation to the evolution of the other strategy. Dynamic models that track the frequencies of alleles underlying multiple reproductive strategies that might influence each others evolution across generations, and that are enacted among relatives that themselves result from reproductive strategies and inbreeding depression expressed in previous generations, are consequently useful to understand and predict evolutionary outcomes.

We use individual-based modelling to address three key questions regarding evolution of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry given opportunity for inbreeding and strong inbreeding depression. First, does evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance facilitate evolution of polyandry that is otherwise severely constrained by increasing relatedness of polyandrous females' additional mates? Second, how do costs associated with polyandry and pre-copulatory and post-copulatory inbreeding avoidance affect evolutionary outcomes and, in particular, the long-term persistence of these reproductive strategies given cost asymmetry? Third, how is selection for initially rare pre-copulatory or post-copulatory inbreeding avoidance affected if the other strategy of inbreeding avoidance is already fixed? **For each question, we isolate the effect of the biological mechanism of interest (i.e., post-copulatory inbreeding avoidance, cost asymmetry, and another fixed strategy of inbreeding avoidance) by running simulations in which the mechanism is present and comparing the results to those of simulations in which the it is absent but all other parameter values are identical.** **We restrict our model to simulate how evolution of traits are affected by their direct costs and indirect benefits due to inbreeding avoidance. Although empirical populations will include additional beneficial traits that might affect evolution of mating strategies, targeting our focus on traits affecting inbreeding and polyandry allows us to better isolate their effects and therefore better develop hypotheses for empirical systems.** By addressing **our three** questions, we illustrate how the simultaneous evolution of multiple interacting degenerate reproductive strategies can generate diverse evolutionary outcomes.

@@

**Complex traits such as reproductive behaviour are likely to be highly polygenic [@Evans2008].** **Hence, in our model,** all individuals have 10 physically unlinked diploid loci, and therefore 20 alleles, underlying each of three reproductive strategy traits (i.e., 30 diploid loci in total): tendency for polyandry ($P_{a}$, i.e., 'polyandry' alleles), pre-copulatory inbreeding strategy ($M_{a}$, i.e., 'mating' alleles), and post-copulatory inbreeding strategy ($F_{a}$, i.e., 'fertilisation' alleles). All alleles can take the value of any real number [i.e., a continuum-of-alleles model; @Kimura1965; @Lande1976; @Reeve2000; @Bocedi2014] and combine additively to determine genotypes ($P_{g}$, $M_{g}$, and $F_{g}$) and resulting phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$) for tendency for polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy, respectively. Each individual's genotypic values $P_{g}$, $M_{g}$, and $F_{g}$ equal the sum of its 20 alleles for each trait. Each individual's phenotypic values for pre-copulatory and post-copulatory inbreeding strategy equal their respective genotypic values ($M_{p} = M_{g}$ and $F_{p} = F_{g}$), where negative and positive values cause inbreeding avoidance and preference, respectively (see details of mating and fertilisation strategies below). **In contrast to inbreeding strategies, the polygenic nature of polyandry cannot translate to continuous phenotypes affecting mate choice [e.g., @Shuker2007; @Evans2013].** **Polyandry is more realistically defined as a "threshold trait", influenced by continuous genetic variation but expressed at some discrete phenotypic value(s) [@Lynch1998; @Roff1996; @Roff1998; @Duthie]. We incorporate this genetic architecture into our model by allowing** phenotypic values for tendency for polyandry **to** equal genotypic values ($P_{p} = P_{g}$) only if $P_{g} \geq 0$, while $P_{p} = 0$ if $P_{g} < 0$ because females cannot choose a negative number of mates. Polyandry is therefore influenced by continuous genetic variation but only expressed when $P_{g} > 0$. A negative $P_{g}$ value therefore generates a female that is phenotypically monandrous, while a positive $P_{g}$ value generates a female that can express some degree of polyandry (see details below). **One important consequence of this genetic architecture is that threshold traits often reappear in a population despite sustained negative selection, as caused by recombination of previously masked deleterious alleles [@Roff1996; @Roff1998].**

In overview, each generation proceeds with females paying costs associated with their reproductive strategy traits, and expressing polyandry, mating, and fertilisation. Offspring inherit a randomly sampled allele from each parent at each locus. Alleles can then mutate and offspring express inbreeding depression in viability. Immigrants arrive in the population and density regulation limits population growth **(see below)**. We record the population pedigree and directly calculate the coefficient of kinship ($k$) between all potential mates in each generation (defined as the probability that two homologous alleles will be identical-by-descent), allowing individual pre-copulatory and post-copulatory inbreeding strategies to be enacted. Key individual traits, parameter values, and variables are described in Table 1.

@@

Phenotypic values of the three reproductive strategy traits each have **pre-determined** costs that are assumed to combine independently to increase the probability that a focal female will die before mating. **The ability to set costs independently and a priori in this way is critical because it permits the direct comparison of evolution of pre-copulatory versus post-copulatory inbreeding avoidance** (see Discussion for consideration **of more mechanistic costs**). The probabilities of pre-mating mortality due to the costs of polyandry ($c_{P}$), pre-copulatory inbreeding strategy ($c_{M}$), and post-copulatory inbreeding strategy ($c_{F}$) are $P_{p} \times c_{P}$, $|M_{p}| \times c_{M}$, and $|F_{p}| \times c_{F}$, respectively. Here $|M_{p}|$ and $|F_{p}|$ are the absolute values of $M_{p}$ and $F_{p}$, respectively. Absolute values are used for applying costs to inbreeding avoidance strategies because both negative and positive $M_{p}$ and $F_{p}$ values could potentially arise and affect the degree of inbreeding, representing inbreeding avoidance and inbreeding preference, respectively. In contrast, $P_{p}$ cannot be negative. **Values of $c_{P}$, $c_{M}$, and $c_{F}$ can be interpreted mechanistically as the increased mortality risk caused by mate searching [e.g., @Koga1998], resisting male harrassment [i.e., sexual conflict; @Rowe1994], and energetic loss from development of post-copulatory inbreeding avoidance mechanisms [e.g., @Gasparini2011], respectively.** Overall, because generations are non-overlapping, a female's probability of total reproductive failure increases linearly with the phenotypic value of each trait.

@@

The probability that an offspring of female $i$ is sired by any one of $i$'s mates $j$ is calculated by assigning a perceived fertilisation quality to each $j$ with respect to $i$, $Q^{f}_{ij}$. Perceived fertilisation quality $Q^{f}_{ij}$ is calculated in the same way as perceived mate quality $Q^{m}_{ij}$, such that if female $i$ has a strategy of post-copulatory inbreeding avoidance ($F_{p}<0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = (-F_{p} \times k_{ij} + 1)^{-1}$. If the female has strategy of post-copulatory inbreeding preference ($F_{p}>0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = F_{p} \times k_{ij} + 1$. A relative quality ($q^{f}_{ij}$) is then calculated for each male by dividing his $Q^{f}_{ij}$ by the sum of the $Q^{f}_{ij}$ values across all of a female's mates. These $q^{f}_{ij}$ values, which lie between zero and one, define the probability of paternity. Females produce $n$ offspring, so a female $i$ samples from her mates $n$ times independently and with replacement with a probability of $q^{f}_{ij}$ for each male $j$ to determine the realised distribution of sires. Offspring have equal probability of being female or male. **After offspring production, all female and male adults die so that generations are non-overlapping.**

@@

The viability of a female $i$'s offspring ($\Psi_{\textrm{off}}$) decreases as a log-linear function of her kinship with the sire $j$ of the offspring ($k_{ij}$) and inbreeding depression slope $\beta$,
\begin{equation}
\Psi_{\textrm{off}} = e^{-\beta k_{ij}}
\end{equation}
Here, $\beta$ models the number of haploid lethal equivalents that exist as deleterious recessive alleles in the gametes of $i$ and $j$, and which might be homozygous in offspring and reduce viability. Equation 1 assumes independent allelic effects, generating multiplicative effects on offspring viability [@Morton1956; @Mills1994]. **It also assumes that inbreeding does not covary with inbreeding load (i.e., no purging), allowing the effects of kinship on inbreeding depression to be consistent across replicate simulations. Additionally, previous modelling has found that inbreeding avoidance in biparental populations has a negligible effect on load [@Duthie2016a].** We model inbreeding depression as having an absolute rather than relative effect on offspring viability (i.e., hard rather than soft selection), so that the effect of $\beta$ is consistent across generations and different parameter combinations. We assume that inbreeding always decreases offspring viability (i.e., $\beta > 0$, giving inbreeding depression but no outbreeding depression). Therefore, because $0 \leq k_{ij} \leq 1$, $-\beta \times k_{ij} \leq 0$. Values of $\Psi_{\textrm{off}}$ must therefore be between zero (if $-\beta \times k_{ij}$ is very negative) and one (if $-\beta \times k_{ij} = 0$). We therefore define $\Psi_{\textrm{off}}$ as the probability that an offspring is viable, and sample its realised viability (versus mortality) using a Bernoulli trial. **Offspring that are viable after the Bernoulli trial become adults.** Given our current objectives, we simulate evolution under conditions where inbreeding avoidance is adaptive due to strong inbreeding depression, and not evolution where inbreeding preference is adaptive due to weak or zero inbreeding depression [@Parker1979; @Kokko2006; @Duthie2016a], or given outbreeding depression [@Bateson1983; @Greeff2009]. However, as described above, positive $M_{p}$ and $F_{p}$ values resulting in inbreeding preference are not precluded from evolving, and could potentially arise due to mutation or drift.

@@

To address how asymmetric costs associated with pre-copulatory and post-copulatory inbreeding avoidance and polyandry affect the long-term persistence of reproductive strategies, we quantify the change in $M_{a}$ and $F_{a}$ over generations in simulations where pre-copulatory inbreeding strategy was cost free ($c_{M}=0$) but post-copulatory inbreeding strategy was moderately costly ($c_{F}=0.02$), and vice versa. We compare evolutionary trajectories with those of a costly strategy in the absence of evolution of an alternative strategy (e.g., evolution of pre-copulatory inbreeding strategy when post-copulatory inbreeding strategy phenotype is fixed at zero, $F_{p}=0$). Previous modelling using similar genetic architecture suggests that a cost of $0.02$ imposes strong but not overwhelming direct negative selection on polyandry [@Duthie]. This value is therefore appropriate to illustrate the different evolutionary consequences that could result from asymmetrical costs **(but see Supporting Information for other asymmetric cost combinations)**.

>>>>

From the perspective of a breeding female, inbreeding avoidance might be achieved by avoiding mating with related males (i.e., pre-copulatory inbreeding avoidance), or by biasing fertilisation towards unrelated males after mating has occurred (i.e., post-copulatory inbreeding avoidance). Evolution of such post-copulatory inbreeding avoidance is likely to be facilitated when females express some degree of polyandry, defined as mating with multiple males during a single reproductive bout [but see @Dougherty2016]. Such polyandry might itself be under indirect selection and evolve if it allows females to mate with additional unrelated males following an initial mating with a relative, or if it allows females to mate with additional unrelated males that would not be available during initial mate choice due to changes in mate availability or relatedness distribution in a population [e.g., @Reid2015; @Duthie]. Polyandry can thereby facilitate pre-copulatory inbreeding avoidance even without any post-copulatory female choice or otherwise biased fertilisation among sperm provided by females' mates (i.e., under conditions of a 'fair raffle').

Overall, polyandry can therefore simultaneously allow females to mate with less closely related males and create the opportunity for further inbreeding avoidance enacted through active post-copulatory choice. Polyandry has consequently been widely hypothesised to allow females to avoid inbreeding and thereby increase offspring fitness [@Zeh1997; @Jennions2000; @Tregenza2002]. Resulting indirect selection on polyandry that stems from the increase in offspring fitness could explain evolution of polyandry in cases where multiple mating does not increase female reproductive success, or even reduces reproductive success and therefore imposes direct costs on polyandrous females. However, despite such widely-invoked hypotheses regarding the role of inbreeding depression in driving pre-copulatory and post-copulatory strategies and polyandry, there is surprisingly little theory or modelling that critically examines the conditions under which pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, are predicted to evolve, or that examines how the evolution of such strategies of inbreeding avoidance might feed back to affect underlying evolution of polyandry. To comprehensively understand evolution of mating strategies given inbreeding depression, it is necessary to consider the fundamental effects of selection on pre-copulatory and post-copulatory inbreeding avoidance and polyandry.

Despite the paucity of theory, numerous empirical studies on diverse species have tested for, and in some cases found evidence of, female inbreeding avoidance in systems where polyandry also occurs [@Tregenza2002; @Varian-Ramos2012; @Kingma2013; @Arct2015; but see @Reid2015a]. However, few studies have determined whether inbreeding avoidance is enacted through pre-copulatory or post-copulatory mechanisms. Pre-copulatory inbreeding avoidance has been reported in sweet potato weevils [*Cylas formicarius*, @Kuriwada2011], purple-crowned fairy-wrens [*Malurus coronatus*, @Kuriwada2011], and squinting bush brown butterflies [*Bicyclus anynana*, @Fischer2015], while post-copulatory inbreeding avoidance has been reported in, e.g., red junglefowl [*Gallus gallus*, @Pizzari2004] and crickets [*Teleogryllus oceanicus*, @Simmons2006; *Gryllus bimaculatus*, @Bretman2009]. Evidence for both pre-copulatory and post-copulatory inbreeding avoidance is available across different studies of Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. Meanwhile, @Liu2014 found evidence of pre-copulatory but not post-copulatory inbreeding avoidance within a single study of cabbage beetles (*Colaphellus bowringi*). However, @Ala-Honkola2011 and @Tan2012 found no evidence for pre-copulatory or post-copulatory inbreeding avoidance in fruit flies (*Drosophila melanogaster*), respectively, and Reid et al. [-@Reid2014; -@Reid2015] found no net inbreeding avoidance in song sparrows (*Melospiza melodia*) despite strong inbreeding depression and opportunities for both pre-copulatory and post-copulatory inbreeding avoidance. Taken together, these studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance, or lack thereof, occur in nature. However, there is as yet no theory that predicts what combinations of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry should be favoured by selection when all can evolve. Consequently, there is no theory that allows the diversity of observed pre-copulatory and post-copulatory inbreeding avoidance strategies to be interpreted, and no clear hypotheses that could be tested through future empirical studies of individual systems or subsequent comparative analyses.

@Duthie used a genetically-explicit individual-based model to examine the conditions under which polyandry is predicted to evolve due to selection stemming from pre-copulatory inbreeding avoidance in the absence of any post-copulatory inbreeding avoidance. Even when selection for pre-copulatory inbreeding avoidance was strong and females consequently preferred unrelated mates, selection for polyandry specifically as a mechanism to facilitate this inbreeding avoidance occurred only under highly restricted conditions, requiring that direct negative selection (i.e., 'costs') on polyandry was weak, that very few males were available for a female's initial mate choice but many were available for additional mate choice(s), or that polyandry was conditionally expressed when a focal female was related to her initial mate [@Duthie]. Without these conditions, females that engaged in polyandry tended to be constrained to increase rather than decrease their overall degree of inbreeding, ultimately reducing offspring fitness. This constraint occurred because, once pre-copulatory inbreeding avoidance evolved, polyandrous females had already chosen available unrelated males as their initial mates. Their additional mates, chosen from the remaining male population, were therefore increasingly likely to include relatives. Evolution of polyandry purely to facilitate pre-copulatory inbreeding avoidance, and hence the resulting degree of inbreeding avoidance, was consequently restricted [@Duthie].

@@

Ultimately, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry will be affected not only by the magnitudes of positive indirect selection stemming from reduced inbreeding depression in females' offspring, but also by the magnitudes of direct negative selection on resulting phenotypes (i.e., the direct fitness costs of expressing each reproductive strategy). There are multiple mechanisms by which direct costs might by incurred for polyandry, pre-copulatory choice, and post-copulatory choice; here we focus on three examples observed in empirical populations. Polyandrous females might pay costs of mate searching or mating because these activities increase predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice and consequent reluctance to mate with relatives might pay costs of increased risk of complete mating failure or harm caused by sexual conflict [@Rowe1994; @Kokko2013]. Finally, females that express post-copulatory choice might pay initial energetic costs associated with developing the physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]. If the relative costs of pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly strategy might become fixed over generations, while alleles underlying the more costly strategy might become extinct, especially if their effects are made redundant by evolution of the less costly strategy. Evolutionary dynamics will therefore depend on the benefits of each strategy, which might also differ and vary in relation to the evolution of the other strategy. Dynamic models that track the frequencies of alleles underlying multiple reproductive strategies that might influence each others evolution across generations, and that are enacted among relatives that themselves result from reproductive strategies and inbreeding depression expressed in previous generations, are consequently useful to understand and predict evolutionary outcomes.

We use individual-based modelling to address three key questions regarding evolution of pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry given opportunity for inbreeding and strong inbreeding depression. First, does evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance facilitate evolution of polyandry that is otherwise severely constrained by increasing relatedness of polyandrous females' additional mates? Second, how do costs associated with polyandry and pre-copulatory and post-copulatory inbreeding avoidance affect evolutionary outcomes and, in particular, the long-term persistence of these reproductive strategies given cost asymmetry? Third, how is selection for initially rare pre-copulatory or post-copulatory inbreeding avoidance affected if the other strategy of inbreeding avoidance is already fixed? For each question, we isolate the effect of the biological mechanism of interest (i.e., post-copulatory inbreeding avoidance, cost asymmetry, and another fixed strategy of inbreeding avoidance) by running simulations in which the mechanism is present and comparing the results to those of simulations in which the it is absent but all other parameter values are identical. We restrict our model to simulate how evolution of traits are affected by their direct costs and indirect benefits due to inbreeding avoidance. Although empirical populations will include additional beneficial traits that might affect evolution of mating strategies, targeting our focus on traits affecting inbreeding and polyandry allows us to better isolate their effects and therefore better develop hypotheses for empirical systems. By addressing our three questions, we illustrate how the simultaneous evolution of multiple interacting degenerate reproductive strategies can generate diverse evolutionary outcomes.

@@

Complex traits such as reproductive behaviour are likely to be highly polygenic [@Evans2008]. Hence, in our model, all individuals have 10 physically unlinked diploid loci, and therefore 20 alleles, underlying each of three reproductive strategy traits (i.e., 30 diploid loci in total): tendency for polyandry ($P_{a}$, i.e., 'polyandry' alleles), pre-copulatory inbreeding strategy ($M_{a}$, i.e., 'mating' alleles), and post-copulatory inbreeding strategy ($F_{a}$, i.e., 'fertilisation' alleles). All alleles can take the value of any real number [i.e., a continuum-of-alleles model; @Kimura1965; @Lande1976; @Reeve2000; @Bocedi2014] and combine additively to determine genotypes ($P_{g}$, $M_{g}$, and $F_{g}$) and resulting phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$) for tendency for polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy, respectively. Each individual's genotypic values $P_{g}$, $M_{g}$, and $F_{g}$ equal the sum of its 20 alleles for each trait. Each individual's phenotypic values for pre-copulatory and post-copulatory inbreeding strategy equal their respective genotypic values ($M_{p} = M_{g}$ and $F_{p} = F_{g}$), where negative and positive values cause inbreeding avoidance and preference, respectively (see details of mating and fertilisation strategies below). In contrast to inbreeding strategies, the polygenic nature of polyandry cannot translate to continuous phenotypes affecting mate choice [e.g., @Shuker2007; @Evans2013]. Polyandry is more realistically defined as a "threshold trait", influenced by continuous genetic variation but expressed at some discrete phenotypic value(s) [@Lynch1998; @Roff1996; @Roff1998; @Duthie]. We incorporate this genetic architecture into our model by allowing phenotypic values for tendency for polyandry to equal genotypic values ($P_{p} = P_{g}$) only if $P_{g} \geq 0$, while $P_{p} = 0$ if $P_{g} < 0$ because females cannot choose a negative number of mates. Polyandry is therefore influenced by continuous genetic variation but only expressed when $P_{g} > 0$. A negative $P_{g}$ value therefore generates a female that is phenotypically monandrous, while a positive $P_{g}$ value generates a female that can express some degree of polyandry (see details below). One important consequence of this genetic architecture is that threshold traits often reappear in a population despite sustained negative selection, as caused by recombination of previously masked deleterious alleles [@Roff1996; @Roff1998].

In overview, each generation proceeds with females paying costs associated with their reproductive strategy traits, and expressing polyandry, mating, and fertilisation. Offspring inherit a randomly sampled allele from each parent at each locus. Alleles can then mutate and offspring express inbreeding depression in viability. Immigrants arrive in the population and density regulation limits population growth (see below). We record the population pedigree and directly calculate the coefficient of kinship ($k$) between all potential mates in each generation (defined as the probability that two homologous alleles will be identical-by-descent), allowing individual pre-copulatory and post-copulatory inbreeding strategies to be enacted. Key individual traits, parameter values, and variables are described in Table 1.

@@

Phenotypic values of the three reproductive strategy traits each have pre-determined costs that are assumed to combine independently to increase the probability that a focal female will die before mating. The ability to set costs independently and a priori in this way is critical because it permits the direct comparison of evolution of pre-copulatory versus post-copulatory inbreeding avoidance (see Discussion for consideration of more mechanistic costs). The probabilities of pre-mating mortality due to the costs of polyandry ($c_{P}$), pre-copulatory inbreeding strategy ($c_{M}$), and post-copulatory inbreeding strategy ($c_{F}$) are $P_{p} \times c_{P}$, $|M_{p}| \times c_{M}$, and $|F_{p}| \times c_{F}$, respectively. Here $|M_{p}|$ and $|F_{p}|$ are the absolute values of $M_{p}$ and $F_{p}$, respectively. Absolute values are used for applying costs to inbreeding avoidance strategies because both negative and positive $M_{p}$ and $F_{p}$ values could potentially arise and affect the degree of inbreeding, representing inbreeding avoidance and inbreeding preference, respectively. In contrast, $P_{p}$ cannot be negative. Values of $c_{P}$, $c_{M}$, and $c_{F}$ can be interpreted mechanistically as the increased mortality risk caused by mate searching [e.g., @Koga1998], resisting male harrassment [i.e., sexual conflict; @Rowe1994], and energetic loss from development of post-copulatory inbreeding avoidance mechanisms [e.g., @Gasparini2011], respectively. Overall, because generations are non-overlapping, a female's probability of total reproductive failure increases linearly with the phenotypic value of each trait.

@@

The probability that an offspring of female $i$ is sired by any one of $i$'s mates $j$ is calculated by assigning a perceived fertilisation quality to each $j$ with respect to $i$, $Q^{f}_{ij}$. Perceived fertilisation quality $Q^{f}_{ij}$ is calculated in the same way as perceived mate quality $Q^{m}_{ij}$, such that if female $i$ has a strategy of post-copulatory inbreeding avoidance ($F_{p}<0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = (-F_{p} \times k_{ij} + 1)^{-1}$. If the female has strategy of post-copulatory inbreeding preference ($F_{p}>0$), then the perceived quality of male $j$ is $Q^{f}_{ij} = F_{p} \times k_{ij} + 1$. A relative quality ($q^{f}_{ij}$) is then calculated for each male by dividing his $Q^{f}_{ij}$ by the sum of the $Q^{f}_{ij}$ values across all of a female's mates. These $q^{f}_{ij}$ values, which lie between zero and one, define the probability of paternity. Females produce $n$ offspring, so a female $i$ samples from her mates $n$ times independently and with replacement with a probability of $q^{f}_{ij}$ for each male $j$ to determine the realised distribution of sires. Offspring have equal probability of being female or male. After offspring production, all female and male adults die so that generations are non-overlapping.

@@

The viability of a female $i$'s offspring ($\Psi_{\textrm{off}}$) decreases as a log-linear function of her kinship with the sire $j$ of the offspring ($k_{ij}$) and inbreeding depression slope $\beta$,
\begin{equation}
\Psi_{\textrm{off}} = e^{-\beta k_{ij}}
\end{equation}
Here, $\beta$ models the number of haploid lethal equivalents that exist as deleterious recessive alleles in the gametes of $i$ and $j$, and which might be homozygous in offspring and reduce viability. Equation 1 assumes independent allelic effects, generating multiplicative effects on offspring viability [@Morton1956; @Mills1994]. It also assumes that inbreeding does not covary with inbreeding load (i.e., no purging), allowing the effects of kinship on inbreeding depression to be consistent across replicate simulations. Additionally, previous modelling has found that inbreeding avoidance in biparental populations has a negligible effect on load [@Duthie2016a]. We model inbreeding depression as having an absolute rather than relative effect on offspring viability (i.e., hard rather than soft selection), so that the effect of $\beta$ is consistent across generations and different parameter combinations. We assume that inbreeding always decreases offspring viability (i.e., $\beta > 0$, giving inbreeding depression but no outbreeding depression). Therefore, because $0 \leq k_{ij} \leq 1$, $-\beta \times k_{ij} \leq 0$. Values of $\Psi_{\textrm{off}}$ must therefore be between zero (if $-\beta \times k_{ij}$ is very negative) and one (if $-\beta \times k_{ij} = 0$). We therefore define $\Psi_{\textrm{off}}$ as the probability that an offspring is viable, and sample its realised viability (versus mortality) using a Bernoulli trial. Offspring that are viable after the Bernoulli trial become adults. Given our current objectives, we simulate evolution under conditions where inbreeding avoidance is adaptive due to strong inbreeding depression, and not evolution where inbreeding preference is adaptive due to weak or zero inbreeding depression [@Parker1979; @Kokko2006; @Duthie2016a], or given outbreeding depression [@Bateson1983; @Greeff2009]. However, as described above, positive $M_{p}$ and $F_{p}$ values resulting in inbreeding preference are not precluded from evolving, and could potentially arise due to mutation or drift.

@@

To address how asymmetric costs associated with pre-copulatory and post-copulatory inbreeding avoidance and polyandry affect the long-term persistence of reproductive strategies, we quantify the change in $M_{a}$ and $F_{a}$ over generations in simulations where pre-copulatory inbreeding strategy was cost free ($c_{M}=0$) but post-copulatory inbreeding strategy was moderately costly ($c_{F}=0.02$), and vice versa. We compare evolutionary trajectories with those of a costly strategy in the absence of evolution of an alternative strategy (e.g., evolution of pre-copulatory inbreeding strategy when post-copulatory inbreeding strategy phenotype is fixed at zero, $F_{p}=0$). Previous modelling using similar genetic architecture suggests that a cost of $0.02$ imposes strong but not overwhelming direct negative selection on polyandry [@Duthie]. This value is therefore appropriate to illustrate the different evolutionary consequences that could result from asymmetrical costs (but see Supporting Information for other asymmetric cost combinations).
